Visual Inference for Networks

### 2016-08-03 JSM/VizInfSlides.Rmd
2 more slides

Goal: model dynamic networks with

- Network effects (e.g. outdegree, reciprocity)
- Actor covariate effects

Observed networks treated as discrete observations of a continuous time Markov Chain.

1. At each (unobserved) time point 1 actor can change 1 tie
2. Between observations all actors are given the opportunity to change at same rate
3. When an actor can change, aims to maximize its objective functon, $f_i(\mathbf{\beta},x) = \sum_k \beta_k s_{ki}(x)$, where $\beta_k$ are network or actor parameters, and $s_{ki}(x)$ are corresponding statistics.

@@

## Model Parameters {.smaller}

@@

```{r echo=FALSE}
dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-rev-m-12-rep-c.csv")
ggplot(data = dat %>% filter(plot_order2 %in% c(11,12)) %>% mutate(plot = plot_order2 - 10) ) + 
  geom_net(aes(from_id = X1, to_id = X2), size = 2, fiteach = T, directed = T, 
           colour = 'black', arrowsize = .25, arrowgap = .02, ecolour = 'grey60') + 
  facet_wrap(~plot) + 
  theme_net()
```

@@

```{r echo=FALSE}
dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-m-12-rep-1.csv")
ggplot(data = dat %>% filter(plot_order %in% c(2,7)) %>% mutate(plot = ifelse(plot_order == 2, 1, 2))) + 
  geom_net(aes(from_id = X1, to_id = X2), size = 2, fiteach = T, directed = T, 
           colour = 'black', arrowsize = .25, arrowgap = .02, ecolour = 'grey60') + 
  facet_wrap(~plot) + 
  theme_net()
```

@@

- Idea: a very statistically significant difference in models should look different to us

@@

- Small Amazon Turk experiment

## Summary

## Significance (?)

# Conclusion

## What we've learned

## Future Work

## References

## Thank you!

>>>>

Observed networks: discrete observations of continuous time Markov Chain.

1. Between observations:  actors act (i.e. change ties with others) at same rate
2. Actor aims to maximize  objective functon, $f_i(\mathbf{\beta},x) = \sum_k \beta_k s_{ki}(x)$.

@@

## Two Models {.smaller}

@@

```{r echo=FALSE}
dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-rev-m-12-rep-c.csv")
ggplot(data = dat %>% filter(plot_order2 %in% c(11,12)) %>% mutate(Model = plot_order2 - 10) ) + 
  geom_net(aes(from_id = X1, to_id = X2), size = 2, fiteach = T, directed = T, 
           colour = 'black', arrowsize = .25, arrowgap = .02, ecolour = 'grey60') + 
  facet_wrap(~Model, labeller = "label_both") + 
  theme_net() + theme(panel.background = element_rect(color = 'black'))
```

@@

```{r echo=FALSE}
dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-m-12-rep-1.csv")
ggplot(data = dat %>% filter(plot_order %in% c(2,7)) %>% mutate(Model = ifelse(plot_order == 2, 2, 1))) + 
  geom_net(aes(from_id = X1, to_id = X2), size = 2, fiteach = T, directed = T, 
           colour = 'black', arrowsize = .25, arrowgap = .02, ecolour = 'grey60') + 
  facet_wrap(~Model, labeller = "label_both") + 
  theme_net() + theme(panel.background = element_rect(color = 'black'))
```

@@

- Idea: a very statistically significant difference in models should look different to us
- Method: visual inference - use lineups with graphs from models 1 & 2 to test this idea

## Lineup

- Show $m$ plots to people
- Ask to identify most different plot
- Null ($m-1$ plots) vs. Alternative ($1$ plot) models
- Identification of the alternative plot is evidence against null
- Probability to pick alternative plot randomly is $\frac{1}{m}$

@@

- M1 v. M2 (most complex structure)
- M2 v. M1 (least complex structure)
- five lineups each
- each participant randomly sees ten lineups

## Most complex?

```{r, echo=FALSE, fig.align='center'}
dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-m-12-rep-10.csv")
ggplot(data = dat) +
  geom_net(aes(from_id = X1, to_id = X2), size = 1, fiteach = T, directed = T, 
           colour = 'black', arrowsize = .15, arrowgap = .02, ecolour = 'grey60') + 
  facet_wrap(~plot_order) + 
  theme_net() + theme(panel.background = element_rect(color = 'black'))
```

## Least Complex?

```{r echo=FALSE, fig.align='center'}
dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-rev-m-12-rep-b.csv")
ggplot(data = dat) +
  geom_net(aes(from_id = X1, to_id = X2), size = 1, fiteach = T, directed = T, 
           colour = 'black', arrowsize = .15, arrowgap = .02, ecolour = 'grey60') + 
  facet_wrap(~plot_order2) + 
  theme_net() + theme(panel.background = element_rect(color = 'black'))
```

## Small Amazon Turk experiment

<img src="screenshot.png" width="800px" height="400px" />

## Results 1

``` {r, echo=FALSE}
tab <- readr::read_csv("../Results/turk21_users.csv")
res <- readr::read_csv("../Results/turk21_feedback.csv")
```

- barchart with filled correct

## Results 2

- scatterplot of %correct by beta 3
- scatterplot of %correct by #jtt

## Conclusion

- Doesn't work for networks
- Need another way! (layout?)
